### 2017-10-16 mcmc.tex
fewer other methods

\author[0000-0002-9328-5652]{Dan Foreman-Mackey}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
\affil{Department~of~Astronomy, University~of~Washington, Box 351580, Seattle, WA 98195, USA}

@@

\paragraph{Gibbs:}
In many problems, the parameters---the components or entries of the
blob $\pars$---are not equivalent at all.
Some are global parameters, which touch or have an effect on every
sub-part of the data blob $\data$, while others are local parameteres,
which only touch a small subset of the data.
Or, in another use case, some parameters are linear parameters
that---at fixed values of the other parameters---will have a Gaussian
likelihood or posterior pdf.
In these problems, there are some directions in parameter space that
are hard to move in (they require complete re-calculation of the full
non-linear function $f(\pars)$, and other directions that are very
easy to move in (they require only partial re-calculations or the
sampling could even be analytic and exact).
For problems with these structures, Gibbs samplers are ideal, and
there are good implementations.\note{For example, \project{JAGS}; find it on the internets.}

@@

Gibbs sampling can be very good for exploiting multiprocessing:
Global parameter updates might require a serial re-calculation of the
full likelihood, but local parameter updates can be done in parallel,
with each local parameter working on its local bit of data all at the
same time.
This kind of structure is common in hierarchical inference, where
local parameters touch only individual objects in some population,
say, and global parameters are parameters \emph{of} that population as
a whole.
Parallelization is built in to some Gibbs implementations.\note{For
  example, \project{JAGS} automatically
  parallelizes Gibbs problems.}

\paragraph{Hamiltonian or Hybrid:}
In many cases of interest, the function $f(\pars)$ can be analytically
differentiated, such that it is possible to quickly execute a gradient
$\dd f/\dd\pars$ evaluation.
Indeed, in the modern world of computing, there are even
auto-differentiation systems that deliver automatically code that
produces the gradient of any function you can write.\note{%
  The idea is that for (almost) any function you can write,
  a robot can write the function that is the derivative of that
  function, using simple differentiation rules and the magic of the
  chain rule. Differentiation is a typographic operation on computer
  functions!  The idea is \emph{not} to take numerical (small
  difference) derivatives of the code; there is no way that
  finite-difference differentiation could help your sampling in the
  long run.}
There is a class of \MCMC\ methods that make use of gradient information
to effectively inform the proposal distribution and thus speed sampling.

DFM: more here!  For example, does this work better in large numbers
of dimensions?  If so, we should say so.  And cite \citep{hamiltonian}.

@@

\paragraph{Slice:}
DFM ... When is slice sampling a good idea?  DFM, can you help here?

Slice sampling makes use of the point that (in one dimension) a fair
sampling of $f(\pars)$ is a projection onto the $\pars$ axis of a
uniform sampling of the two-dimensional region under the
two-dimensional plot of $f(\pars)$ \vs\ $\pars$.
The $D$-dimensional generalization is that a fair sampling of
$f(\pars)$ is a projection onto the $\pars$ $D$-dimensional
hypersurface of $\pars$ of a uniform sampling of the $D+1$-dimensional
hypervolume under the hypersurface $f(\pars)$ in a hypothetical plot
of $f(\pars)$ over the whole parameter space.
Thinking in terms of the two-dimensional case (for simplicity), slice
samplers make both ``horizontal'' moves in the $\pars$ direction at
fixed $f(\pars)$ and also ``vertical'' moves in the $f(\pars)$
direction.
Because the sampling in the higher-dimensional space is uniform, it is
easy to make moves to very different parts of parameter space, or
different peaks or modes in the function $f(\pars)$.
In our own research, we use a variant of slice sampling for problems
involving samplings governed by Gaussian Processes.\note{DFM: Cite us
  and Murray here.}

@@

\paragraph{Data augmentation:}
...HOGG Look at the relevant chapter in Brooks.

>>>>

\author[0000-0002-9328-5652]{Dan Foreman-Mackey}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
\affil{Department~of~Astronomy, University~of~Washington, Box 351580, Seattle, WA 98195, USA}

@@

\paragraph{Gibbs:}
Neither \MH\ \MCMC\ nor most ensemble methods good at going to very
  large numbers of parameters.
Even dozens of parameters can be enough to drive autocorrelation times
  extremely large.
With larger numbers of parameters, new methods need to be considered.
Gibbs sampling is worth considering when the number of parameters is
  large, and the different parameters have different ``scopes'' in the problem.

In many problems that have large numbers of parameters,
  the parameters---the components or entries of the
  blob $\pars$---are not equivalent at all.
Some are global parameters, which touch or have an effect on every
sub-part of the data blob $\data$, while others are local parameteres,
which only touch a small subset of the data.
Or, in another use case, some parameters are linear parameters
that---at fixed values of the other parameters---will have a Gaussian
likelihood or posterior pdf.
In these problems, there are some directions in parameter space that
are hard to move in (they require complete re-calculation of the full
non-linear function $f(\pars)$, and other directions that are very
easy to move in (they require only partial re-calculations or the
sampling could even be analytic and exact).
For problems with these structures, Gibbs samplers are ideal, and
there are good implementations.\note{For example, \project{STAN} and \project{JAGS}; find them on the internets. DFM INSERT URLs AND REFERENCES.}

@@

Gibbs sampling can also be very good for exploiting multiprocessing:
Global parameter updates might require a serial re-calculation of the
full likelihood, but local parameter updates can be done in parallel,
with each local parameter working on its local bit of data all at the
same time.
This kind of structure is common in hierarchical inference, where
local parameters touch only individual objects in some population,
say, and global parameters are parameters \emph{of} that population as
a whole.
Parallelization is built in to some Gibbs implementations.\note{For
  example, \project{JAGS} automatically
  parallelizes Gibbs problems.}

\paragraph{Hamiltonian:}
At time of writing, the premier approaches for problems with large numbers
  of parameters are Hamiltonian methods.\note{For a review of the theory
  behind these methods, see \cite{hamiltonian}.}
These methods perform very well at large dimensions, but require (for speed)
  analytic derivatives of the function $f(\pars)$ with respect to the parameters.

In many cases of interest, the function $f(\pars)$ can be analytically
differentiated, such that it is possible to quickly execute a gradient
$\dd f/\dd\pars$ evaluation.
Indeed, in the modern world of computing, there are even
auto-differentiation systems that deliver automatically code that
produces the gradient of any function you can write.\note{%
  The idea is that for (almost) any function you can write,
  a robot can write the function that is the derivative of that
  function, using simple differentiation rules and the magic of the
  chain rule. Differentiation is a typographic operation on computer
  functions!  The idea is \emph{not} to take numerical (small
  difference) derivatives of the code; there is no way that
  finite-difference differentiation could help your sampling in the
  long run.}
There is a class of \MCMC\ methods that make use of gradient information
to effectively inform the proposal distribution and thus speed sampling.

DFM: more here! The reason these methods are ``Hamiltonian'' is that... And cite \citep{hamiltonian}.

@@

@@
